\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{geometry}
\geometry{margin=1in}

\title{Unified Simple Field Theory: Particle--Universe Report}
\author{Generated Summary}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This report summarizes the ``Unified Simple Field Theory'' particle--universe sandbox implemented in \texttt{universe\_sim.py}.
The model is intentionally compact and assumption-closed: certainty gradients drive effective gravity, spin phase drives torsion, a two-sheet coupling approximates matter/backsheet interactions, and a horizon drift term induces an outward flow.
It is a configurable toy model for intuition-building, not a claim of general relativity or Einstein--Cartan dynamics.
\end{abstract}

\section{State Variables and Dimensions}
Each particle carries:
\begin{itemize}
  \item A $6$-dimensional position $x_i \in \mathbb{R}^6$ and velocity $v_i \in \mathbb{R}^6$.
  \item A scalar uncertainty $\sigma_i$ (bounded by floor/ceiling clamps).
  \item A spin phase $s_i \in [-\pi,\pi)$.
  \item A sheet label (matter vs backsheet) that flips cross-sheet coupling.
\end{itemize}
Diagnostics are visualized using a 2D projection of the full state, representing a single observer's view.

\section{Core Definitions}
Certainty and heat are defined as
\begin{align}
  C_i &= \frac{1}{\sigma_i^2},\\
  H_i &= \frac{1}{C_i} = \sigma_i^2.
\end{align}
A buoyancy factor modulates interaction strength and sign,
\begin{equation}
  b_i = 1 - \frac{\sigma_i}{\sigma_{\mathrm{crit}}}.
\end{equation}
Large $\sigma_i$ (low certainty) yields negative or reduced coupling, while small $\sigma_i$ strengthens attraction.

\section{Forces and Dynamics}
The net acceleration on particle $i$ is a sum of:
(i) certainty-gradient gravity,
(ii) spin torsion,
(iii) optional ``black-hole'' core enhancement,
and (iv) mild velocity damping for stability.

\subsection{Certainty-Gradient Gravity}
For each pair $(i,j)$, define a sheet coupling factor
\begin{equation}
  s_{ij} =
  \begin{cases}
    1, & \text{same sheet},\\
    -\beta, & \text{cross-sheet},
  \end{cases}
\end{equation}
where $\beta \in (0,1]$ is the reduced backsheet coupling.
The pair weight is
\begin{equation}
  w_{ij} = G_c\,(C_i - C_j)\,b_i b_j\,s_{ij}\,g_{\mathrm{bh}}(j),
\end{equation}
with $g_{\mathrm{bh}}(j)$ boosting attraction when particle $j$ is a high-certainty ``core'' (\,$\sigma_j < \sigma_{\mathrm{bh}}$\,).
The resulting radial acceleration is proportional to
\begin{equation}
  \mathbf{a}_i^{(\mathrm{grav})} = -\sum_{j\neq i} w_{ij}\,\frac{\mathbf{r}_{ij}}{\lVert \mathbf{r}_{ij} \rVert^3},
\end{equation}
where $\mathbf{r}_{ij} = x_i - x_j$.

\subsection{Spin Torsion}
Spin differences generate a transverse force in the projected $(x,y)$ plane:
\begin{equation}
  \Delta s_{ij} = s_i - s_j, \qquad
  \mathbf{a}_i^{\perp} \propto \sum_{j\neq i} \sin(\Delta s_{ij})\,\frac{\mathbf{r}_{ij}^{\perp}}{\lVert \mathbf{r}_{ij} \rVert + \epsilon}.
\end{equation}
A smooth core profile limits the short-range divergence, and the same sheet-coupling sign convention applies.

\subsection{Horizon Drift}
An outward terminal flow models an entropy sink at radius $R$:
\begin{equation}
  \mathbf{v}_{\mathrm{drift}} = v_0\,\frac{y}{(1-y)^{\alpha}}\,\hat{r},
  \qquad y = \frac{r}{R}.
\end{equation}
This induces a dark-energy-like expansion near the boundary.

\section{Thermodynamics of Uncertainty}
Uncertainty evolves via cooling toward a floor and heating from kinetic energy:
\begin{equation}
  \dot{\sigma}_i = -\lambda\,(\sigma_i - \sigma_{\mathrm{floor}}) + \eta\,\lVert v_i \rVert^2,
\end{equation}
with clamping to $[\sigma_{\mathrm{floor}},\sigma_{\mathrm{ceiling}}]$.
In this closure, cold regions can ``crystallize'' to low uncertainty while energetic regions heat and diffuse.

\section{Two-Sheet Fabric and Singular Cores}
The model includes a backsheet (antimatter) that evolves with a time-reversed sign convention and reduced cross-sheet coupling.
Particles that cool below $\sigma_{\mathrm{bh}}$ behave like singular cores with enhanced attraction and transverse frame-dragging.

\section{Diagnostics}
Optional diagnostics track mean kinetic energy, mean certainty/heat, and RMS radius.
Visualization includes particle positions, radial and tangential speed trends, and the uncertainty distribution histogram.

\section{Notes and Intended Use}
This report documents the sandbox closure as implemented.
The goal is a compact, editable model in which each term can be adjusted independently for experimentation.
It is not a claim of physical completeness.

\end{document}
